---
title: "Assignment 2"
author: "Junfu Huang_27494152; Bocong Zhao_28305329"
date: "23/09/2020"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r initial, echo = FALSE, cache = FALSE, results = 'hide'}
knitr::opts_chunk$set(echo = FALSE,warning=FALSE)
```

```{r, message=FALSE, include = FALSE}
library(tidyverse)
library(ozmaps)
library(ggplot2)
library(sf)
library(ggpubr)
```

# Introduction 

**Australia is a suitable place for animals and there are abundant species. People raise different animals and live peacefully with wild animals. However, a certain number of animals are not treated properly, RSPCA which is the largest animal welfare organization in Australia receives these animals and solves problems every year. The dataset is collected through RSPCA's statistics on animal outcomes in every state and year. Based on the dataset, our report provides an analysis of the change in the number of each animals' outcomes for different states from 1999 to 2018.**

```{r , message=FALSE, include = FALSE}
tuesdata <- tidytuesdayR::tt_load(2020, week = 30)
animal <- tuesdata$animal_outcomes
```

# Data Analysis 

## figure 1

```{r}
animal %>% select(-Total,-outcome) %>%
  pivot_longer(cols = -c('animal_type','year'),names_to='State',values_to='value') %>%
  ggplot(aes(x=animal_type,y=value)) +
  geom_boxplot()+ 
  facet_wrap(~State,ncol=4,scales="free_y")+ scale_y_log10()+ 
   theme(axis.text.x = element_text(angle = 40)) +ylab("log(value)")+xlab("Animal Type") + ggtitle("Figure 1:The boxplot of animal in different State")
```

**Figure 1 portrays the distribution of the total number of different animal types per year each outcome in each state. Combining with summary statistics, there is a huge gap between different animal sizes, therefore, we choose to transform the value through log y-axis in order to reduce the effect of variance. From figure 1, we find that most states would have the highest distribution and median value of cat or dog (25% to 75%). From the boxplot, even though we have transformed the value, it still has outliers in the plot, such as dog's values. It means that there is small number of dogs in some years in some states. **

**Because people are more likely to adopt cats or dogs, Vic and NSW are the two most populous states, that is why Victoria and New South Wales have the highest value. By contrast, for example, Western Australia has many desserts, and is sparsely populated. It might have less value for dog or cat than VIC or NSW. Apart from the cat or dog, QLD has large distribution for wildlife. It might be because the living environment is good for wildlife, such as lizards, kangaroos.**

## Figure 2

**Based on figure 1, we choose data in year 2018 as an example.**

```{r , message=FALSE, include = FALSE}
# Calculate Dog numbers in Year 2018
Dog <- animal %>% 
  filter(animal_type=="Dogs", year == "2018") # %>% 
  #na.omit()
colSums(Dog[c("NSW", "VIC", "QLD","SA","WA","TAS", "NT", "ACT")])
DogData <- c(372, 7381, 11517 , 2442, 821 , 452, 10075, 803 ,0)
sf_oz<- ozmap_data("states")
sf_oz$DogData <- DogData

# Calculate Cat numbers in Year 2018
Cat <- animal %>% 
  filter(animal_type=="Cats", year == "2018") # %>% 
  #na.omit()
colSums(Cat[c("NSW", "VIC", "QLD","SA","WA","TAS", "NT", "ACT")])
CatData <- c(452, 12272, 13648, 5070, 940, 1325, 15688, 1775,0)
sf_oz1<- ozmap_data("states")
sf_oz1$CatData <- CatData

# Calculate Horses numbers in Year 2018
Horses <- animal %>% 
  filter(animal_type=="Horses", year == "2018") # %>% 
  #na.omit()
colSums(Horses[c("NSW", "VIC", "QLD","SA","WA","TAS", "NT", "ACT")])
HorseData <- c(0, 107, 125, 27, 40, 7, 173, 0 ,0)
sf_oz2<- ozmap_data("states")
sf_oz2$HorseData <- HorseData

# Calculate Livestock numbers in Year 2018
Livestock <- animal %>% 
  filter(animal_type=="Livestock", year == "2018") # %>% 
  #na.omit()
colSums(Livestock[c("NSW", "VIC", "QLD","SA","WA","TAS", "NT", "ACT")])
LivestockData <- c(9, 326, 1597, 194, 169, 86, 847, 81,0)
sf_oz3<- ozmap_data("states")
sf_oz3$LivestockData <- LivestockData

# Calculate Other Animals numbers in Year 2018
Other_Animals <- animal %>% 
  filter(animal_type=="Other Animals", year == "2018") # %>% 
  #na.omit()
colSums(Other_Animals[c("NSW", "VIC", "QLD","SA","WA","TAS", "NT", "ACT")])
OtherAnimalData <- c(10,1096,2236,590,191,145,2380,494,0 )
sf_oz4<- ozmap_data("states")
sf_oz4$OtherAnimalData <- OtherAnimalData

# Calculate Wildlife numbers in Year 2018
Wildlife <- animal %>% 
  filter(animal_type=="Wildlife", year == "2018") # %>% 
  #na.omit()
colSums(Wildlife[c("NSW", "VIC", "QLD","SA","WA","TAS", "NT", "ACT")])
WildlifeData <- c(0, 1112, 25385, 267, 22, 19, 1038, 340 ,0 )
sf_oz5<- ozmap_data("states")
sf_oz5$WildlifeData <- WildlifeData
```

```{r}
#sf_oz$NAME

P1<- ggplot(data=sf_oz, aes(fill=DogData))+
  geom_sf()+
  scale_fill_gradient(low="green", high = "red")+
  labs(title ="Dog Amounts")+
  theme_classic()+
  theme(axis.line=element_blank(), 
        axis.text = element_blank(),
        axis.ticks = element_blank())

P2<-ggplot(data=sf_oz, aes(fill=CatData))+
  geom_sf()+
  scale_fill_gradient(low="green", high = "red")+
  labs(title = "Cat Amounts")+
  theme_classic()+
  theme(axis.line=element_blank(), 
        axis.text = element_blank(),
        axis.ticks = element_blank())

P3<- ggplot(data=sf_oz, aes(fill=HorseData))+
  geom_sf()+
  scale_fill_gradient(low="green", high = "red")+
  labs(title ="Horse Amounts")+
  theme_classic()+
  theme(axis.line=element_blank(), 
        axis.text = element_blank(),
        axis.ticks = element_blank())

P4<- ggplot(data=sf_oz, aes(fill=LivestockData))+
  geom_sf()+
  scale_fill_gradient(low="green", high = "red")+
  labs(title ="Livestock Amounts")+
  theme_classic()+
  theme(axis.line=element_blank(), 
        axis.text = element_blank(),
        axis.ticks = element_blank())

P5<-ggplot(data=sf_oz, aes(fill=OtherAnimalData))+
  geom_sf()+
  scale_fill_gradient(low="green", high = "red")+
  labs(title = "Other Animal Amounts")+
  theme_classic()+
  theme(axis.line=element_blank(), 
        axis.text = element_blank(),
        axis.ticks = element_blank())

P6<-ggplot(data=sf_oz, aes(fill=WildlifeData))+
  geom_sf()+
  scale_fill_gradient(low="green", high = "red")+
 labs(title = "Wildlife Amounts")+
  theme_classic()+
  theme(axis.line=element_blank(), 
        axis.text = element_blank(),
        axis.ticks = element_blank())

figure <- ggarrange(P1, P2, P3,P4,P5,P6,
                   # labels = c("Dog", "Cat", "Horse", "Livestock","Other Animal","Wildlife"),
                    ncol = 2, nrow = 3)
figure
```

**The map shows the number of species in each Australian state in 2018 in the form of visual graphics, states with relatively small numbers of animals appearing in green and relatively large numbers of animals in red. Showing the number of different animals on the map of Australia can help readers understand, explore and compare the "relationships" between each state.**

**Looking vertically to compare the amount of different types of animals within the same state in 2018. Overall, Queensland has the largest amounts of animals for dogs, livestock and wildlife. The Northern Territory has the largest amounts of animals for cats, horses, and other animals. New South Wales has relatively smaller numbers for all types of animals. As for wildlife, except Queensland, other states have similar situations.**

**Looking horizontally to compare the scale of different types of animals among different states in 2018.  Horse has the smallest scale indicated overall it has the smallest number compared to other types of animals in every state. While wildlife has the largest scale indicated overall it has the largest number compared to other types of animals in every state.**

## Figure 3

```{r}
animal %>%select(-(4:11)) %>% ggplot(aes(x=year,y=Total,col=outcome)) +
  geom_line() +
  facet_wrap(~animal_type,scales = 'free_y') + ylab('The number of animal') +
  ggtitle('Figure 3: The different outcomes in different year for each animal')
```

**Figure 3 shows for each animal, the different outcome over the years. For cats: the number of accepting euthanasia is fluctuated increasing from 1999 to 2007. However, the number of euthanasia has dropped dramatically after 2007. At the same time, the number of rehomed is gradually rising. people might be more likely to rehome cats compared other outcomes. For the dogs: in the past 20 years, the number of dogs undergoing euthanasia or rehome is gradually decreasing. The rehomed or reclaimed dogs still account for a high number. For the horses, after 2007, horses might remain in the shelter for care. For the Livestock and other animals, they may be more likely to be rehomed compared with other methods, especially for the other animals (like rabbits), the number of rehomed is continuously increasing. However, the organizations receive wildlife, they might choose euthanasia wildlife. And the number of euthanasia has been rising after 2011.**

**The RSPCA may choose to help dogs or cats to find the new owners rather than let them die. Conversely, when RSPCA received the wildlife, it means that wildlife may be difficult to survive due to injury or other reasons, the RSPCA might choose euthanasia.  **

## Figure 4

```{r}
animal %>% group_by(animal_type,year)%>%summarise(to=sum(Total)) %>% ggplot(aes(x=year,y=to)) +
  geom_point() + geom_smooth() +
  facet_wrap(~animal_type,scales = 'free_y')+ ylab('The number of animal')+
  ggtitle('Figure 4: The trend of animal for the year')
```

**The plots show year against total values, faceted by animal type. From the data description, animal types include cats, dogs, horses, livestock, wildlife and other animals. Thus, the graph shows the overall variation tendency for each different animal through years.**

**Cats and dogs fluctuate on a similar scale, around 0 to 70000. The number of cats showed a sharp increase from year 2000 to year 2007 with the amount increased from 48798 to 65380. Then it shows an obvious decreasing pattern afterwards with the amount decreasing to 51688. From year 2015 onwards, it appears a slowly increasing pattern with the amount increased to 53554.**

**There is a tiny decrease in the number of dogs between year 1999 to year 2001 with amounts between 67204 to 63613. Then it shows a slowly increasing pattern until year 2008 with the amount up to 69383. Finally, it suddenly appears a sharp decrease afterwards with the amount to only 365-6.**

**The amount of horses and other animals both appear to have an overall sharp increasing pattern. The difference is that the number of horses show an overall increasing pattern, even though there are some fluctuations in between. The amount is 118 in year 1999 and at the final collected data in year 2018, the amount has increased to 516. However, as the scale is between 0 to 600, horses still have the smallest amount among all types of animals. While, the number of other animals show an obviously sharp increasing pattern until year 2012, however, it shows a slowly decreasing pattern afterwards.**

**The amount of livestock does not have an obviously changing pattern compared to other types of animals.  There is a stable tiny increasing pattern before year 2006 and a stable tiny decreasing pattern after year 2011. Within 2006 and 2011 is the bloom period for the amount of livestock, with the amount increased from 2842 to 3691.**

**The amount of wildlife has a relatively stable pattern until 2007. Then it appears an obviously increasing pattern afterwards with the amount increased from 11973 to 30151.**

# Limitation and Conclusion 

**The dataset only contains the number of rehomed, reclaimed and euthanized varioused by animal type, states and years. But it does not give any information about the factors which influence people to make the decisions about rehomed, reclaimed and euthanized. This maybe due to people's psychological factors which are hard to measured and reflect on data.**

**More specific, our analysis finds cats and dogs are easier to rehomed or reclaimed, but the number of euthanized is still not small. Therefore, we may want to know what kinds of cats or dogs are more likely to rehomed, reclaimed, and why cats or dogs would be euthanized. This maybe due to people's attitudes to dogs and cats' characteristics are very unique and hard to give a common summary and statistic. Or even lack conditions such as technique, labor, time to collect such information.**

**On the other hand, Australia is rich in wildlife species. The dataset does not classify the type of wild animals. As the range of wild animals is very board which indicates the bodily form and species would varied by type. And further influence the number of rehomed or reclaimed. However, the report could not provide evidence for why wildlife is euthanized based on the dataset.**

**Talking about future behavior and advice. As from figure 4, there is an increasing trend of wild animals. So we might need to consider that apart from euthanized, can we do better for them. Also, there is an decreasing pattern for dogs and cats, maybe indicates that people are love their pets more and less discard them.**

# Appendix & Reference 

Thomas Mock, Georgios Karamanis (2020). Australian Pets: Animal Outcomes  https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-21/readme.md

```{r}
animal$animal_type <-as.factor(animal$animal_type)
animal$outcome <- as.factor(animal$outcome)
animal%>% summary()
```
